---
title: "Result Report: Label Effects on Perceptions and Behaviors Toward Unhoused Individuals"
# If blank, the running header is the title in upper case.
shorttitle: "Label Effects and Homlessness"
# Set names and affiliations.
# It is nice to specify everyone's orcid, if possible.
# There can be only one corresponding author, but declaring one is optional.
author:
  - name: Huidi Yuan
    affiliation: "University of Chicago"
abstract: "This study investigates how inclusive labels, compared to stigmatizing terms, shape public perceptions and prosocial behaviors toward unhoused individuals. Drawing on theories of labeling and stigma, we hypothesized that person-centered labels (e.g., “people experiencing housing insecurity”) would reduce stigma and stereotypes (H1) and increase donations (H2). A preregistered online experiment with 400 U.S. adults tested these hypotheses using a between-subjects design. Participants viewed materials referencing either “the homeless” (stigmatizing label) or “people experiencing housing insecurity” (inclusive label). Results supported H1: PC labels significantly reduced stigma and negative stereotypes. However, H2 was unsupported, as donations did not differ between conditions. Stigma negatively correlated with donation amounts, but stereotypes showed no relationship. Exploratory analyses found no moderation by political orientation or age. These findings highlight the potential of PC labels to mitigate stigma, though additional strategies are needed to bridge the gap between attitudinal and behavioral change. Implications for advocacy, policy communication, and future research are discussed."
# Put as many keywords at you like, separated by commmas (e.g., [reliability, validity, generalizability])
keywords: [Inclusive Language, stigma, homelessness, prosocial behavior, labeling effects]
# Suppress author note
suppress-author-note: true
# If true, tables and figures are mingled with the text instead of listed at the end of the document.
floatsintext: true
# Numbered lines (.pdf and .docx only)
numbered-lines: false
# File with references
bibliography: PC_Label_Homeless_Reference.bib
# Suppress title page
suppress-title-page: false
# Link citations to references
link-citations: true
# If true, adds today's date below author affiliations. If text, can be any value.
# This is not standard APA format, but it is convenient.
# Works with docx, html, and typst. 
draft-date: false
# Language options. See https://quarto.org/docs/authoring/language.html
lang: en
format:
  apaquarto-docx: default
  apaquarto-html: default
  apaquarto-typst: 
    keep-typ: true
  apaquarto-pdf:
    # Can be jou (journal), man (manuscript), stu (student), or doc (document)
    documentmode: man
    keep-tex: true
execute:
  echo: false # set global chunk options: Show output only
  warning: false # override these in individual chunks
---
# Introduction

The way individuals are labeled—such as "the homeless" versus "people experiencing housing insecurity"—can influence stigma, stereotypes, and charitable behavior. This study examines how different labels affect:

- Perceptions of homelessness-related stigma,
- Stereotypical judgments,
- Donation behaviors.

We hypothesize that:

1. Person-centered labels reduce stigma and stereotypes (H1); 
2. Person-centered labels increase donations to unhoused group (H2);
3. Stigma and stereotypes predict donation amounts to unhoused group (H3).

## Methods

```{r}
#| label: setup
#| include: false

# Source library R script
source("06_New_Codes/library.R")

```

```{r}
#| label: data-load
#| include: false

# Read data
data_deidentify <- read.csv("data_deidentify.csv")

# Move ppt_ID to the first column
data_deidentify <- data_deidentify %>%
  select(ppt_ID, everything())

view(data_deidentify)
```

```{r}
#| label: cleaning-separate-demo
#| include: false

# Separate demographics
data_demo <- data_deidentify %>%
  select(ppt_ID, 
         label,
         starts_with("Demo_"), 
         comment
         )

write.csv(data_demo, "data_demo.csv", row.names = FALSE)
```

```{r}
#| label: cleaning-separate-fr
#| include: false

# Separate free response questions
data_fr <- data_deidentify %>%
  select(ppt_ID, 
         label,
         general_1,
         general_2,
         like_fr,
         offens_fr,
         impression_1, 
         impression_2,
         impression_3_4_TEXT,
         comment
         )
```

```{r}
#| label: cleaning-separate-quant-main
#| include: false

# Separate the MAIN DATA that is not demo or free response
data_main <- data_deidentify %>%
  select(-c(starts_with("Demo_"),
            general_1,
            general_2,
            like_fr,
            offens_fr,
            impression_1,
            impression_2,
            impression_3_4_TEXT,
            comment)
         ) %>%
  # Correct data type
  mutate(across(where(is.character), # Apply to character columns only
                ~ ifelse(!is.na(suppressWarnings(as.numeric(.))),
                         as.numeric(.), 
                         .))) %>% # If it can be converted to numeric, convert it
  # Correct data type for columns with NA
  mutate(
    experience_1 = as.numeric(experience_1),
    experience_2 = as.numeric(experience_2),
    experience_3 = as.numeric(experience_3),
    SliderOverlapValue = as.numeric(SliderOverlapValue)
  ) %>%
  # Select rows that pass attention check
  filter(check == 2)
  
```

```{r}
#| label: cleaning-recode
#| include: false

# Recode to correct wrong values and wrong order
data_main_recode <- data_main %>%
  # In primary and early_language, change 4=English and other to 3 = English and other.
  mutate(
    primary = if_else(primary == 4, 3, primary),
    early_language = if_else(early_language == 4, 3, early_language)
  ) %>%
  # For all stigma_, 5 = the most stigmatized, 1 = the least stigmatized
  ## Recode stigma_resp_2, stigma_dist_2, stigma_dist_3
  mutate(stigma_resp_2_mut = 6 - stigma_resp_2,
         stigma_dist_2_mut = 6 - stigma_dist_2, 
         stigma_dist_3_mut = 6 - stigma_dist_3) %>%
  select(-c(stigma_resp_2, stigma_dist_2, stigma_dist_3)) %>%
  # For all stertyp_, 5 = the most stereotyped (negatively)
  ## Reverse all values in stertyp_ columns
  mutate(across(
    starts_with("stertyp_"), # Select columns starting with "stertyp_"
    ~ 6 - ., # Reverse the values 
    .names = "{.col}_rev" # Add "_rev" to the new column names
  )) %>%
  ## Remove original stertyp_ columns: start with "stertyp_" but not end with "_rev"
  select(-starts_with("stertyp_")|ends_with("_rev")) %>%
  # Calculate main DVs: average stigma and stereotype
  mutate(
    stigma_avg = rowMeans(select(., starts_with("stigma_")), na.rm = TRUE),
    stereotype_avg = rowMeans(select(., starts_with("stertyp_")), na.rm = TRUE)
  ) %>%
  # Rename donate columns
  rename(donate_unhoused = donate_1, 
         donate_children = donate_2,
         donate_animals = donate_3)

```

### Participants
A total of `r nrow(data_deidentify)` participants were recruited. `r nrow(data_deidentify) - nrow(data_main_recode)` participants were excluded due to failing attention checks. Thus the final sample size was `r nrow(data_main_recode)`.

### Measures
- **Stigma Perception** (6-item 5-points Likert scale)
- **Stereotypical Judgment** (8-item 5-points Likert scale) [@fiskeetal_2002_model]
- **Donation Allocation** (Amount donated to homeless organizations out of three different organizations)

|                  | **Measures**  | **Items**                             |
|------------------|---------------|---------------------------------------|
| | **Stigma**  _N = 6_ | [Label] are responsible for their condition  |
| | _Likert (1-5)_      | [Label] are dangerous  |
| **Survey Measures** | | It is better to be away from [Label].  |
| | **Stereotypes**  _N = 8_ | How competent are [Label]?  |
| | _Likert (1-5)_           | How well educated are [Label]?  |
| |                          | How warmhearted are [Label]?  |
| **Behavioral Measures** | **Donation** _$100_ | Covenant House - For [Label]  |
||                                                  | Save the Children  |
|                     |                                                  | American Society for the Prevention of Cruelty to Animals (ASPCA)  |

: Main Measures {#tbl-measures apa-note="Stigma measures include: Responsibility, danger, social distance; Stereotypes measures include: Competence, warmth, status, competition."
data-quarto-disable-processing="true"}

# Result
## Person-Centered Labels Reduce Stigma and Stereotypes (H1)
### Stigma
```{r}
#| label: f-test-stigma-by-label
#| include: false

# F-test (ANOVA) for stigma by label
stigma_aov <- aov(stigma_avg ~ label, data = data_main_recode)
summary(stigma_aov)

```
Stigma perception significantly differed by label.

```{r}
#| label: fig-stigma-by-label
#| caption: "Stigma by Label"

# Boxplot of stigma by label
ggplot(data_main_recode, aes(x = label, y = stigma_avg, fill = label)) +
  geom_boxplot() +
  labs(title = "Boxplot of Average Stigma Perception by Label",
    x = "Label",
    y = "Average Stigma Perception") +
  theme_minimal() +
  theme(legend.position = "none") +
  stat_compare_means(method = "anova",
                     label.x = 1.3,
                     label.y = 5)
```

### Stereotypes
```{r}
#| label: f-test-stereotype-by-label
#| include: false

# F-test (ANOVA) for stereotype by label
stereotype_aov <- aov(stereotype_avg ~ label, data = data_main_recode)
summary(stereotype_aov)
```
Stereotype perception significantly differed by label.

```{r}
#| label: fig-stereotype-by-label
#| caption: "Stereotypes by Label"

# Boxplot of stereotypes by label
ggplot(data_main_recode, aes(x = label, y = stereotype_avg, fill = label)) +
  geom_boxplot() +
  labs(title = "Boxplot of Average Stereotype Perception by Label",
    x = "Label",
    y = "Average Stereotype Perception") +
  theme_minimal() +
  theme(legend.position = "none") +
  stat_compare_means(method = "anova",
                     label.x = 1.3,
                     label.y = 5)
```

### Stigma and Stereotypes Correlation
```{r}
#| label: correlation-stigma-stereotype
#| include: false

# Correlation between stigma and stereotypes
cor_stigma_stereotype <- cor.test(data_main_recode$stigma_avg, data_main_recode$stereotype_avg, method = "pearson")

cor_stigma_stereotype
```
Stigma and stereotypes were significantly correlated.

```{r}
#| label: fig-correlation-stigma-stereotype
#| caption: "Correlation between Stigma and Stereotypes"

# Scatterplot and linear regression of stigma and stereotypes
ggplot(data_main_recode, aes(x = stigma_avg, y = stereotype_avg, color = label)) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(title = "Correlation between Stigma and Stereotypes",
    x = "Average Stigma Perception",
    y = "Average Stereotype Perception") +
  theme_minimal() +
  theme(legend.position = "none")
```

## Person-Centered Labels Increase Donations (H2)
### Donation Distribution
```{r}
#| label: donation-distribution-by-label
#| include: false

# Mean donate_1,2,3 by label
mean_donate <- data_main_recode %>%
  group_by(label) %>%
  summarise(
    mean_donate_unhoused = mean(donate_unhoused, na.rm = TRUE),
    mean_donate_children = mean(donate_children, na.rm = TRUE),
    mean_donate_animals = mean(donate_animals, na.rm = TRUE)
  )
```
Among the three organizations, avergae donation to the unhoused organization is the highest (report mean donation to unhoused) compared to children (report mean donation to children) and animal (report mean donation to animal) organizations. See figure... 

### Donation to Unhoused Org. by Label
```{r}
#| label: f-test-donation-by-label
#| include: false

# F-test (ANOVA) for donation by label
supportive_aov <- aov(donate_unhoused ~ label, data = data_main_recode)

summary(supportive_aov)
```

Donation, however, did not significantly differ by label.
[report mean donation for unhoused org. by label]


```{r}
#| label: fig-donate-all-by-label
#| caption: "Donation Distribution by Label"

# Pivot long for donate_1,2,3
data_main_recode_donate_long <- data_main_recode %>%
  pivot_longer(cols = starts_with("donate"), names_to = "organization", values_to = "donation")

# Boxplot of donate by organization and label
ggplot(data_main_recode_donate_long, aes(x = organization, y = donation, fill = label)) +
  geom_boxplot() +
  labs(title = "Boxplot of Donation Amount by Organization and Label",
    x = "Organization",
    y = "Donation Amount") +
  theme_minimal() +
  theme(legend.position = "none") +
  stat_compare_means(method = "anova",
                     label.x = 1.3,
                     label.y = 105)
```

## Stigma and stereotypes predict donation (H3)
### Stigma and Donation
```{r}
#| label: correlation-stigma-donation
#| include: false

# Pearson correlation between stigma and donate_unhoused
cor_stigma_donate <- cor.test(data_main_recode$stigma_avg, data_main_recode$donate_unhoused, method = "pearson")

cor_stigma_donate
```

```{r}
#| label: correlation-stigma-donation-by-label
#| include: false

# Hierarchical Linear Regression for stigma predicting donate_unhoused by label

## Center stigma_avg
data_main_recode_scale_stigma <- data_main_recode %>%
  mutate(stigma_avg_c = scale(stigma_avg, center = TRUE, scale = FALSE))

## Run Moderation Analysis
moderation_model <- lm(donate_unhoused ~ stigma_avg_c * label, data = data_main_recode_scale_stigma)

## Display results
summary(moderation_model)
```
Stigma perception is negatively correlated with donation amount. However, the relationship between stigma and donation did not differ by label.

```{r}
#| label: fig-correlation-stigma-donation
#| caption: "Correlation between Stigma and Donation to Unhoused Organization"

# Scatterplot and linear regression of stigma and donate_unhoused
ggplot(data_main_recode, aes(x = stigma_avg, y = donate_unhoused, color = label)) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(title = "Correlation between Stigma and Donation to Unhoused Organization",
    x = "Average Stigma Perception",
    y = "Donation Amount") +
  theme_minimal() +
  theme(legend.position = "none")
```

### Stereotypes and Donation
```{r}
#| label: correlation-stereotype-donation
#| include: false

# Pearson correlation between stereotype and donate_unhoused
cor_stereotype_donate <- cor.test(data_main_recode$stereotype_avg, data_main_recode$donate_unhoused, method = "pearson")

cor_stereotype_donate
```
Stereotype perception is not significantly correlated with donation amount.


\newpage
# References
::: {#refs}
:::