---
title: "PC_English_Full"
output:
  pdf_document: default
  html_document: default
date: "2024-11-18"
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Cleaning

```{r load library}
library(readr)
library(tidyverse)
library(ggplot2)
library(ggsignif)
library(mlogit)
```

```{r read original data}
data <- read_csv("99_Original_Data/PC_Pilot-English_Prolific_December 12, 2024_17.20.csv")
View(data)
```

```{r remove header}
data_header_rm <- data[-c(1,2),] #remove qualtrics duplicate headers
data_header_rm[data_header_rm == ""] <- NA #empty string = NA
View(data_header_rm)
``` 

```{r consent and finish}
# Prolific, pass consent, and prescreen
data_passed <- data_header_rm %>%
  filter(
    Finished == 1,
    consent == 1,
    str_length(Prolific_ID) == 24, # Prolific IDs
  )
```

```{r remove fraud} 
#rows with fraud
data_fraud <- data_passed %>%
  filter(
    Q_BallotBoxStuffing == "True",
    Q_RecaptchaScore <= 0.5,
    Q_RelevantIDDuplicate == "True",
    Q_RelevantIDDuplicateScore >= 75,
    Q_RelevantIDFraudScore >= 30
  )

#remove unnecessary columns
data_passed1 <- data_passed %>%
  filter(!PROLIFIC_PID %in% data_fraud$PROLIFIC_PID) %>% #remove fraud
  select(-c(StartDate,
            EndDate,
            Status,
           IPAddress,
           Progress,
           Finished,
           RecordedDate,
           ResponseId,
           RecipientLastName,
           RecipientFirstName,
           RecipientEmail,
           ExternalReference,
           LocationLatitude,
           LocationLongitude,
           DistributionChannel,
           UserLanguage,
           'Duration (in seconds)',
           Q_BallotBoxStuffing,
           Q_RecaptchaScore,
           Q_RelevantIDDuplicate,
           Q_RelevantIDDuplicateScore,
           Q_RelevantIDFraudScore,
           PROLIFIC_PID,
           consent, 
           not_pc_label, 
           pc_label
           )) %>%
  mutate(ppt_ID = row_number())
```

```{r}
# Separate demographics
data_demo <- data_passed1 %>%
  select(ppt_ID, Prolific_ID,
         label,
         starts_with("Demo_"), 
         comment
         )

write.csv(data_demo, "demo_12102024.csv", row.names = FALSE)

# Main data
data_main <- data_passed1 %>%
  select(-c(Prolific_ID, # de-identified
            starts_with("Demo_"), 
            comment)) %>%
  mutate(across(where(is.character), # Apply to character columns only
                ~ ifelse(!is.na(suppressWarnings(as.numeric(.))),
                         as.numeric(.), 
                         .))) %>%
  mutate(
    experience_1 = as.numeric(experience_1),
    experience_2 = as.numeric(experience_2),
    experience_3 = as.numeric(experience_3),
    SliderOverlapValue = as.numeric(SliderOverlapValue)
  )

# Attention check
data_main <- data_main[data_main$check == 2, ]

write.csv(data_main, "data_clean.csv", row.names = FALSE)

# Main data with demo
data_main_with_demo <- data_passed1 %>%
  select(-c(Prolific_ID, # de-identified
            comment)) %>%
  mutate(across(where(is.character), # Apply to character columns only
                ~ ifelse(!is.na(suppressWarnings(as.numeric(.))),
                         as.numeric(.), 
                         .))) %>%
  mutate(
    experience_1 = as.numeric(experience_1),
    experience_2 = as.numeric(experience_2),
    experience_3 = as.numeric(experience_3),
    SliderOverlapValue = as.numeric(SliderOverlapValue)
  )

# Attention check
data_main_with_demo <- data_main_with_demo[data_main_with_demo$check == 2, ]

write.csv(data_main_with_demo, "data_clean_with_demo.csv", row.names = FALSE)

```

## Recode Data

```{r}
# In primary and early_language, change 4=English and other to 3 = English and other. 
data_main <- data_main %>%
  mutate(
    primary = if_else(primary == 4, 3, primary),
    early_language = if_else(early_language == 4, 3, early_language)
  )

# For all stigma_, 5 = the most stigmatized, 1 = the least stigmatized
## Recode stigma_resp_2, stigma_dist_2, stigma_dist_3
data_main_recode <- data_main %>%
  mutate(stigma_resp_2_mut = 6 - stigma_resp_2,
         stigma_dist_2_mut = 6 - stigma_dist_2, 
         stigma_dist_3_mut = 6 - stigma_dist_3) %>%
  select(-c(stigma_resp_2, stigma_dist_2, stigma_dist_3))

# For all stertyp_, 5 = the most stereotyped (negatively)
## Reverse all values in stertyp_ columns
data_main_recode <- data_main_recode %>%
  mutate(across(
    starts_with("stertyp_"), # Select columns starting with "stertyp_"
    ~ 6 - ., # Reverse the values (assuming range is 1 to 5)
    .names = "{.col}_rev" # Add "_rev" to the new column names
  )) %>%
  select(-starts_with("stertyp_")|ends_with("_rev"))

data_main_noFR <- data_main_recode %>%
  select(-c(general_1, general_2, impression_1, impression_2,impression_3_4_TEXT, like_fr, offens_fr))
write.csv(data_main_noFR, "data_main_noFR.csv", row.names = FALSE)

```

## Analysis - Simple Descriptive

### Lable meaning overlap

```{r}
# Calculate stigma and stereotype as averages
data_main_recode <- data_main_recode %>%
  mutate(
    stigma_avg = rowMeans(select(., starts_with("stigma_")), na.rm = TRUE),
    stereotype_avg = rowMeans(select(., starts_with("stertyp_")), na.rm = TRUE)
  )

# Overlap compare
data_main_recode %>%
  group_by(label) %>%
  summarise(overlap_avg = mean(SliderOverlapValue, na.rm = TRUE))

# Overlap vs. stigma correlation
cor_overlap_stigma <- cor.test(data_main_recode$SliderOverlapValue, data_main_recode$stigma_avg, method = "pearson")

cor_overlap_stigma

# Overlap vs. stereotype correlation
cor_overlap_stereotype <- cor.test(data_main_recode$SliderOverlapValue, data_main_recode$stereotype_avg, method = "pearson")

cor_overlap_stereotype
```

### Stigma: PC vs. PIC

```{r}
# average stigma 
data_main_recode %>%
  mutate(stigma_avg_ppt = rowSums(select(., starts_with('stigma_')))/6) %>% 
  group_by(label) %>%
  summarise(stigma_avg_label = mean(stigma_avg_ppt))
  
# stigma_resp
data_main_recode %>%
  mutate(stigma_resp_ppt = rowSums(select(., starts_with('stigma_resp')))/2) %>% 
  group_by(label) %>%
  summarise(stigma_resp_label = mean(stigma_resp_ppt))

# stigma_dist
data_main_recode %>%
  mutate(stigma_dist_ppt = rowSums(select(., starts_with('stigma_dist')))/3) %>% 
  group_by(label) %>%
  summarise(stigma_dist_label = mean(stigma_dist_ppt))

# stigma_danger
data_main_recode %>%
  group_by(label) %>%
  summarise(stigma_danger_label = mean(stigma_danger))
```

### Stereotype: PC vs. PIC

```{r}
# average stertyp
data_main_recode %>%
  mutate(stertyp_avg_ppt = rowSums(select(., starts_with('stertyp_')))/8) %>% 
  group_by(label) %>%
  summarise(stertyp_avg_label = mean(stertyp_avg_ppt))

# stertyp_cmptition
data_main_recode %>%
  mutate(stertyp_cmptition_ppt = rowSums(select(., starts_with('stertyp_cmptition_')))/2) %>% 
  group_by(label) %>%
  summarise(stertyp_cmptition_label = mean(stertyp_cmptition_ppt))

# stertyp_comp
data_main_recode %>%
  mutate(stertyp_comp_ppt = rowSums(select(., starts_with('stertyp_comp_')))/2) %>% 
  group_by(label) %>%
  summarise(stertyp_comp_label = mean(stertyp_comp_ppt))

# stertyp_warm
data_main_recode %>%
  mutate(stertyp_warm_ppt = rowSums(select(., starts_with('stertyp_warm_')))/2) %>% 
  group_by(label) %>%
  summarise(stertyp_warm_label = mean(stertyp_warm_ppt))

# stertyp_status
data_main_recode %>%
  mutate(stertyp_status_ppt = rowSums(select(., starts_with('stertyp_status_')))/2) %>% 
  group_by(label) %>%
  summarise(stertyp_status_label = mean(stertyp_status_ppt))
```

### Supportive Behavior: PC vs. PIC

```{r}
# Average donation: PC vs. PIC
data_main_recode %>%
  group_by(label) %>%
  summarise(donate_homeless_label = mean(donate_1), donate_homeless_sd = sd(donate_1))

ggplot(data_main_recode, aes(x = donate_1)) + 
  geom_histogram() + 
  facet_wrap(~label)
  
ggplot(data_main_recode, aes(x = donate_2)) + 
  geom_histogram() + 
  facet_wrap(~label)
  
ggplot(data_main_recode, aes(x = donate_3)) + 
  geom_histogram() + 
  facet_wrap(~label)
```

### Concept: PC vs. PIC

```{r}
# concept_1: access to shelter
data_main_recode %>%
  group_by(label) %>%
  summarise(concept_shelter_lable = mean(concept_1))

# concept_2: unemployment
data_main_recode %>%
  group_by(label) %>%
  summarise(concept_unemploy_lable = mean(concept_2))

# concept_3: duration of homelessness
data_main_recode %>%
  group_by(label) %>%
  summarise(concept_duration_lable = mean(concept_3))
```

### Like better & Less offensive: PC vs. PIC

```{r}
# Average like, offens: if distinct from 1.5 
data_main_recode %>%
  summarise(like_avg_ppt = mean(like), 
            offense_avg_ppt = mean(offens))

## like, offens: if correlated with label
data_main_recode %>%
  group_by(label) %>%
  summarise(like_avg_label = mean(like), 
            offense_avg_label = mean(offens))
```

### Impression - Which one is more politically correct

```{r}
# Distribution of impression_3
data_main_recode %>%
  summarise(impression_3_avg = mean(impression_3))

data_main_recode$impression_3 <- factor(data_main$impression_3, 
                            levels = c(1, 2, 3, 4), 
                            labels = c("PIC", "PC", "Both", "Neither"))

ggplot(data_main_recode, aes(x = impression_3)) +
  geom_bar() +
  labs(title = "Distribution of Impression_3", 
       x = "Impression_2", 
       y = "Frequency") +
  theme_minimal()

data_main_recode %>%
  summarise(familiar_1_avg_ppt = mean(familiar_1), 
            familiar_2_avg_ppt = mean(familiar_2))
```

## Analysis - F Test

```{r}


# Perform F-tests for each variable
# 1. Stigma
stigma_aov <- aov(stigma_avg ~ label, data = data_main_recode)
summary(stigma_aov)

## plot stigma by label
summary_stigma_aov <- data_main_recode %>%
  group_by(label) %>%
  summarise(
    stigma_mean = mean(stigma_avg), 
    stigma_se = sd(stigma_avg)/sqrt(n())
  )

ggplot(summary_stigma_aov, aes(x = label, y = stigma_mean, fill = label)) + 
  geom_bar(stat = "identity", position = "dodge") + 
  geom_errorbar(aes(ymin = stigma_mean - stigma_se, 
                    ymax = stigma_mean + stigma_se), 
                width = 0.2) +
  geom_text(aes(label = round(stigma_mean, 2), y = stigma_mean), vjust = -0.5) +
  labs(title = "Bar Graph of Stigma Averages by Label",
    x = "Label",
    y = "Average Stigma Score") +
  ylim(0, 5) +
  scale_fill_manual(
    values = c("#B0B0B0", "#404040")  # Custom colors for the fill aesthetic
  ) +
  theme_minimal() +
  theme(legend.position = "none")

# 2. Stereotype
stereotype_aov <- aov(stereotype_avg ~ label, data = data_main_recode)
summary(stereotype_aov)

## plot stertyp by label
summary_stereotype_aov <- data_main_recode %>%
  group_by(label) %>%
  summarise(
    stereotype_mean = mean(stereotype_avg), 
    stereotype_se = sd(stereotype_avg)/sqrt(n())
  )

ggplot(summary_stereotype_aov, aes(x = label, y = stereotype_mean, fill = label)) + 
  geom_bar(stat = "identity", position = "dodge") + 
  geom_errorbar(aes(ymin = stereotype_mean - stereotype_se, 
                    ymax = stereotype_mean + stereotype_se), 
                width = 0.2) +
  geom_text(aes(label = round(stereotype_mean, 2), y = stereotype_mean), vjust = -0.5) +
  labs(title = "Bar Graph of Stereotype Averages by Label",
    x = "Label",
    y = "Average Stereotype Score") +
  ylim(0, 5) +
  scale_fill_manual(
    values = c("#B0B0B0", "#404040")  # Custom colors for the fill aesthetic
  ) +
  theme_minimal() +
  theme(legend.position = "none")
  

# 3. Supportive Behavior
supportive_aov <- aov(donate_1 ~ label, data = data_main_recode)
summary(supportive_aov)

## plot donate_1 by label
summary_donate_aov <- data_main_recode %>%
  group_by(label) %>%
  summarise(
    donate_mean = mean(donate_1), 
    donate_se = sd(donate_1)/sqrt(n())
  )

ggplot(summary_donate_aov, aes(x = label, y = donate_mean, fill = label)) + 
  geom_bar(stat = "identity", position = "dodge") + 
  geom_errorbar(aes(ymin = donate_mean - donate_se, 
                    ymax = donate_mean + donate_se), 
                width = 0.2) +
  geom_text(aes(label = round(donate_mean, 2), y = donate_mean), vjust = -1.5) +
  labs(title = "Bar Graph of Donation by Label",
    x = "Label",
    y = "Average Donation") +
  scale_fill_manual(
    values = c("#B0B0B0", "#404040")  # Custom colors for the fill aesthetic
  ) +
  theme_minimal() +
  theme(legend.position = "none")
```

Significant difference in stigma (F(1, 389) = 8.245, p = 0.0043) and stereotype (F(1, 389) = 7.433, p = 0.0067) between PC and PIC. No significant difference in donation (F(1, 389) = 0.705, p = 0.402).

## Analysis - Correlation (Pearson)

### Correlation between stigma and stereotype

```{r}
# Pearson correlation between stigma and stereotype
cor_stigma_stereotype <- cor.test(data_main_recode$stigma_avg, data_main_recode$stereotype_avg, method = "pearson")

cor_stigma_stereotype
```

Stigma and stereotype are positively correlated (r = 0.45, p \< 0.001).

### Correlation between stigma and donate_1

```{r}
# Pearson correlation between stigma and donate_1
cor_stigma_donate <- cor.test(data_main_recode$stigma_avg, data_main_recode$donate_1, method = "pearson")

cor_stigma_donate
```

### Correlation between stereotype and donate_1

```{r}
# Pearson correlation between stereotype and donate_1
cor_stereotype_donate <- cor.test(data_main_recode$stereotype_avg, data_main_recode$donate_1, method = "pearson")

cor_stereotype_donate
```

```{r}

```

## Analysis - Multinomial Logistic Regression

```{r}
# Model 1: Stata

# Model 2: Cleaning 
## political orientation
data_main_recode <- data_main_recode %>%
  mutate(political = rowMeans(select(., c(political_social, political_econ)), na.rm = TRUE))

## label preference
### like: 1 = the homeless/PIC, 2 = people who experience housing insecurity/PC

## experience: 1 = have personal or close experience, 0 = don't have experience
data_main_recode <- data_main_recode %>%
  mutate(experience = ifelse(experience_1 == 1 | experience_2 == 1, 1, 0))

# Replace NA with 0 in 'experience'
data_main_recode$experience[is.na(data_main_recode$experience)] <- 0

## empathy
data_main_recode <- data_main_recode %>%
  mutate(empathy = rowMeans(select(., starts_with("empathy_")), na.rm = TRUE))


data_main_noFR <- data_main_recode %>%
  select(-c(general_1, general_2, impression_1, impression_2,impression_3_4_TEXT, like_fr, offens_fr))
write.csv(data_main_noFR, "data_main_noFR.csv", row.names = FALSE)

```

## Exploratory Analysis

### Distribution among three organizations (donate_1, donate_2, donate_3)

```{r}
data_main_recode %>%
  group_by(label) %>%
  summarise(donate_1_mean = mean(donate_1), donate_2_mean = mean(donate_2), donate_3_mean = mean(donate_3))

# Bar graph of donation by label
summary_donate_all <- data_main_recode %>%
  group_by(label) %>%
  summarise(
    mean_1 = mean(donate_1), 
    se_1 = sd(donate_1)/sqrt(n()), 
    mean_2 = mean(donate_2),
    se_2 = sd(donate_2)/sqrt(n()),
    mean_3 = mean(donate_3),
    se_3 = sd(donate_3)/sqrt(n())
  )

summary_donate_all_long <- summary_donate_all %>%
  pivot_longer(
    cols =  mean_1:se_3,  # Columns to reshape
    names_to = c("metric", "donate_type"),  # Split names into two parts
    names_sep = "_",  # Separator between "mean" and type
    values_to = "value"  # New column for the reshaped values
  ) %>%
  pivot_wider(
    names_from = metric,  # Spread 'mean' and 'se' into columns
    values_from = value
  )

ggplot(summary_donate_all_long, aes(x = donate_type, y = mean, fill = label)) +
  geom_bar(stat = "identity", position = "dodge", width = 0.7) +
  geom_errorbar(aes(ymin = mean - se, 
                    ymax = mean + se), 
                position = position_dodge(0.7),
                width = 0.2) +
  geom_text(aes(label = round(mean, 2), y = mean), 
            position = position_dodge(0.7), 
            vjust = -1.5) +
  labs(
    title = "Comparison of Mean Donations to Each Org. by Label",
    x = "Donation to Organization",
    y = "Mean Donation",
    fill = "Label"
  ) +
  scale_fill_manual(
    values = c("#B0B0B0", "#404040")  # Custom colors for the fill aesthetic
  ) +
  theme_minimal() +
  theme(legend.position = "none")
```

### Political Orientation
```{r}
# political_soical distribution
ggplot(data_main_recode, aes(x = political_social)) +
  geom_bar()

# Recode political_social
data_main_politic_re <- data_main_recode %>%
  filter(political_social != 4) %>%
  mutate(political_social_con = if_else(political_social < 4, "conservative", "liberal"))

# Average donation by political_social
data_main_politic_re %>% group_by(label,political_social_con) %>%
  summarise(donate_1_mean = mean(donate_1))

# political_econ distribution
ggplot(data_main_recode, aes(x = political_econ)) +
  geom_bar()

# Recode political_econ
data_main_politic_re <- data_main_politic_re %>%
  filter(political_econ != 4) %>%
  mutate(political_econ_con = if_else(political_econ < 4, "conservative", "liberal"))

# Average donation by political_econ
data_main_politic_re %>% group_by(label,political_econ_con) %>%
  summarise(donate_1_mean = mean(donate_1))

# Overlap by political_social
data_main_politic_re %>% group_by(political_social_con) %>%
  summarise(overlap_avg = mean(SliderOverlapValue, na.rm = TRUE))

# t test for overlap by political_social
overlap_politic_t <- t.test(SliderOverlapValue ~ political_social_con, data = data_main_politic_re)

overlap_politic_t

# preference by political_social
data_main_politic_re %>% group_by(label,political_social_con) %>%
  summarise(like_avg = mean(like))

data_main_politic_re <- data_main_politic_re %>%
  mutate(like = ifelse(like == 1, 0, 1))

politic_like_lm <- glm(like ~ political_social_con + label, family = "binomial", data = data_main_politic_re)
summary(politic_like_lm)

# prefer by label
prefer_like_aov <- aov(like ~ label, data = data_main_politic_re)
summary(prefer_like_aov)

data_main_politic_re %>% group_by(label) %>%
  summarise(like_avg = mean(like))

data_main_politic_re %>% group_by(political_social_con) %>%
  summarise(like_avg = mean(like))

politic_like_lm <- glm(like ~ political_social_con + label + political_social_con*label, family = "binomial", data = data_main_politic_re)
summary(politic_like_lm)
```

Between conservative and liberal, is there difference in the effect of label on how much they donate?

```{r}
# 2(label)x2(political_social_con) ANOVA for donate_1
politic_social_aov <- aov(donate_1 ~ label * political_social_con, data = data_main_politic_re)

summary(politic_social_aov)

# Effect of political_social_con on stigma
politic_stigma_lm <- lm(stigma_avg ~ political_social_con, data = data_main_politic_re)
summary(politic_stigma_lm)

politic_stigma_t <- t.test(stigma_avg ~ political_social_con, data = data_main_politic_re)
print(politic_stigma_t)

# Effect of political_social_con on donate_1
politic_donate_lm <- lm(donate_1 ~ political_social_con, data = data_main_politic_re)
summary(politic_donate_lm)

politic_donate_t <- t.test(donate_1 ~ political_social_con, data = data_main_politic_re)
print(politic_donate_t)
```

There is no significant interaction effect between label and political_social_con on donate_1 (F(1, 389) = 0.137, p = 0.712).

Plot the political orientation effect on donation by label
```{r}
summary_donate_politic_aov <- data_main_politic_re %>%
  group_by(label, political_social_con) %>%
  summarise(
    donate_mean = mean(donate_1), 
    donate_se = sd(donate_1)/sqrt(n())
  ) %>%
  mutate(label = ifelse(label == "the homeless", "PIC", "PC"))

ggplot(summary_donate_politic_aov, aes(x = political_social_con, y = donate_mean, fill = label)) +
  geom_bar(stat = "identity", position = "dodge", width = 0.7) +
  geom_errorbar(aes(ymin = donate_mean - donate_se, 
                    ymax = donate_mean + donate_se), 
                position = position_dodge(0.7),
                width = 0.2) +
  geom_text(aes(label = round(donate_mean, 2), y = donate_mean), 
            vjust = -2.5,
            position = position_dodge(0.7)) +
  labs(title = "Donation by Label and Political Orientation",
    x = "Political Orientation",
    y = "Average Donation") +
  scale_fill_manual(
    values = c("#B0B0B0", "#404040")  # Custom colors for the fill aesthetic
  ) +
  theme_minimal()
```

Check: Label difference within liberals
```{r}
ggplot(data_main_politic_re, aes(x = label, fill = political_social_con)) +
  geom_bar()

data_main_politic_re_liberal <- data_main_politic_re %>%
  filter(political_social_con == "liberal")

supportive_aov_liberal <- aov(donate_1 ~ label, data = data_main_politic_re_liberal)
summary(supportive_aov_liberal)
```

### Political Orientation and Stigma
```{r}
# Stigma by political_social
data_main_politic_re %>% group_by(political_social_con) %>%
  summarise(stigma_avg = mean(stigma_avg))

stigma_politic_aov <- aov(stigma_avg ~ political_social_con, data = data_main_politic_re)
summary(stigma_politic_aov)

# Stigma by label and political_social
data_main_politic_re %>% group_by(label,political_social_con) %>%
  summarise(stigma_avg = mean(stigma_avg))

stigma_label_politic_aov <- aov(stigma_avg ~ label * political_social_con, data = data_main_politic_re)

summary(stigma_label_politic_aov)

# Plot stigma by label and political_social
summary_stigma_label_politic_aov <- data_main_politic_re %>%
  group_by(label, political_social_con) %>%
  summarise(
    stigma_mean = mean(stigma_avg), 
    stigma_se = sd(stigma_avg)/sqrt(n())
  ) %>%
  mutate(label = ifelse(label == "the homeless", "PIC", "PC"))

ggplot(summary_stigma_label_politic_aov, aes(x = political_social_con, y = stigma_mean, fill = label)) +
  geom_bar(stat = "identity", position = "dodge", width = 0.7) +
  geom_errorbar(aes(ymin = stigma_mean - stigma_se, 
                    ymax = stigma_mean + stigma_se), 
                position = position_dodge(0.7),
                width = 0.2) +
  geom_text(aes(label = round(stigma_mean, 2), y = stigma_mean), 
            vjust = -2,
            position = position_dodge(0.7)) +
  labs(title = "Stigma by Label and Political Orientation",
    x = "Political Orientation",
    y = "Average Stigma") +
  scale_fill_manual(
    values = c("#B0B0B0", "#404040")  # Custom colors for the fill aesthetic
  ) +
  theme_minimal()
```
There are main effects of label and political_social on stigma. But there is no significant interaction effect between label and political_social on stigma.

### Age
```{r}
data_main_political_demo <- left_join(data_main_politic_re, data_demo, by = "ppt_ID") 

data_main_political_demo$demo_age <- as.numeric(data_main_political_demo$demo_age)

# Age distribution
ggplot(data_main_political_demo, aes(x = demo_age)) +
  geom_histogram(stat = "count")

# age by political orientation
data_main_political_demo %>% group_by(political_social_con) %>%
  summarise(demo_age_avg = mean(demo_age))

# if age significantly different by political orientation
age_politic_t <- t.test(demo_age ~ political_social_con, data = data_main_political_demo)
age_politic_t

# donation correlation with age
donate_age_cor <- cor.test(data_main_political_demo$donate_1, data_main_political_demo$demo_age, method = "pearson")

donate_age_cor

# Linear regression graph of donation by age
ggplot(data_main_political_demo, aes(x = demo_age, y = donate_1)) +
  geom_point() +
  geom_smooth(method = "lm", se = TRUE) +
  labs(title = "Donation by Age",
       x = "Age",
       y = "Donation")

# predict donation by age, label, and political_social using
donate_age_label_politic_lm <- lm(donate_1 ~ demo_age + label.x + political_social_con, data = data_main_political_demo)
summary(donate_age_label_politic_lm)

# predict donation by age, label, and age*label
donate_age_label_politic_lm <- lm(donate_1 ~ demo_age + label.x + label.x*demo_age, data = data_main_political_demo)
summary(donate_age_label_politic_lm)

# Pearson correlation between age and donation by political_social_con
cor_age_donate_by_poli <- data_main_political_demo %>%
  group_by(label.x) %>%
  summarise(cor_age_donate = cor(demo_age, donate_1, method = "pearson"))

age_donate



# stigma correlation with age
stigma_age_cor <- cor.test(data_main_political_demo$stigma_avg, data_main_political_demo$demo_age, method = "pearson")

stigma_age_cor

# Pearson correlation between age and stigma by label
cor_age_stigma_by_label <- data_main_political_demo %>%
  group_by(label.x) %>%
  summarise(cor_age_stigma = cor(demo_age, stigma_avg, method = "pearson"))

cor_age_stigma_by_label

# Run Moderation Analysis
moderation_model_stereotype <- lm(stigma_avg ~ demo_age * label.x, data = data_main_political_demo)

summary(moderation_model_stereotype)

# Linear regression graph of stigma versus age by label
ggplot(data_main_political_demo, aes(x = demo_age, y = stigma_avg, color = label.x)) +
  geom_point() +
  geom_smooth(method = "lm", se = TRUE) +
  labs(title = "Stigma by Age and Label",
       x = "Age",
       y = "Stigma") +
  theme_minimal() +
  theme(legend.position = "bottom")


# predict stigma by age, label, and political_social
stigma_age_label_politic_lm <- lm(stigma_avg ~ demo_age + label.x + political_social_con, data = data_main_political_demo)
summary(stigma_age_label_politic_lm)

# predict stigma by age, label, and age*label
stigma_age_label_politic_lm <- lm(stigma_avg ~ demo_age + label.x + label.x*demo_age, data = data_main_political_demo)
summary(stigma_age_label_politic_lm)
```

```{r}
# Within people who think “housing insecurity” is less offensive, compare how many people who like “housing insecurity” and like “the homeless”. 
data_main_political_demo %>% group_by(offens_fr) %>%
  summarise(like_avg = (like))

# how many people think "housing insecurity" is less offensive
data_main_recode %>% group_by(like, offens) %>%
  summarise(n = n())
```

## Next Steps

literature: How does **stigma** perception of one group affect the **behavior** towards this group?

example questions (stigma)
plot results by label
stigma by political 
description of donation 
donation by label 

Q: what should we do with German 
sample can be liberal
should we change behavioral measure? If so, how about feild experiment? Sign up for newsletter? 

# Thesis
different labels exists
stigma may be influenced by label. 
stigma may influecne donation behaviors. 

# Next steps: 
Overlap predicts stigma? overlap dif between label? 

decrease sitgma toward perople in the group? 
Could pity be the pathway? 
gonation = willingness to give something
how much should a city spend to adrress homeless problem/help them? 
homelss vs. unhoused vs. rooflessness 
political vs. preference

move donation to the beginning? 

change study before conducting German study 

run country where homeless is not a big issue? can we find the proper phrase to refer to homeless? 

motivation: what is most effective label to use to generate the most donation? how to build a good branding for homeless organization? 


```{r Qualitative data}
# select columns for qualitative data
data_qual <- data_main_recode %>%
  select(ppt_ID, general_1, general_2, like_fr, offens_fr, impression_1, impression_2)

# write to csv
write.csv(data_qual, "PC_label_qual.csv", row.names = FALSE)
```


